Cartogram Mapping and its Application to Cancer Data Visualization

### 2019-09-25 review_article.Rmd

Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display. Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM]. The use of these supports were found in various online cancer maps identified by @roberts2019communication. The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls. For example, changing the population age or other demographic variables. Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.

The interactivity of modern mapping methods enables supplementary information to be incorporated without cluttering the screen.
Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction]. Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).

These interactive features provide an opportunity for users to explore the additional information available.
Tool tip features and drop down menus allow for user exploration, small screens are not a problem with the flexibility provided by zooming and panning [@HTLWM]. This allows relationships between spatial areas and diseases to be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA].This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution. This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA]. The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.

<!--

Technological advances may provide a solution to this through the use of interactivity, animation, and sound as additional channels for representing uncertainty information without interfering with the ability to see any features that are present in the display [@maceachren1992visualizing]. However, if uncertainty information is perceived as integral and inseparable from the data it is associated with, assumptions of uncertainty creating cognitive overload may not be as valid.

@@

Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map as it changes due to the physical actions of users. 
The manipulation of the web display may occur by directly altering the display, clicking the controls or command-line prompts [@RPIG].
@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
This does not require direct user interaction on the display, but may allow users to start, pause or end the narrative displayed.

@@

<!-- Interactivity (Compare movies or tv to video games) -->
The progress from static print media displays to interactive methods allows users to develop maps to ask and answer their own questions. The affects of these actions are rapid, incremental, and reversible @DMIV. Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM].

Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory. Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA]. Tool tips allow users to obtain additional information regarding specific geographic areas.

Development and adoption of this technology has allowed map users more control of the displayed message. Figure \ref{fig:interacting} presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables. 
These controls may also change the geographies shown, such as continents or countries, states or counties. These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
Interactions with online atlases can also change the amount of detail shown.
It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data.
Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points. 
Low level interaction is more common, this occurs via the user interface or the map.
Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.

@@

Animations can be implemented to allow users varying degrees of control. 
A basic form of interactivity with animations is the play, pause and rewind of films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
Where the message is most important, static or animated graphics allow control over the display and interpretation.
User interaction more allow clicks to progress between stages of more complex animations.
The Australian Cancer Atlas [-@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their own exploration.

@@

>>>>

@@

Interactive methods involve direct manipulation of web displays, keeping a continuous view of the map as it changes due to the physical actions of users. 
The manipulation of the web display may occur by directly altering the display, clicking the controls or command-line prompts [@RPIG].
@TGA explains that animations show the transitions between frames in a linear, fixed narrative to passive users, communicating a message by capturing and directing their attention. 
This may require direct user interaction on the display.

@@

Interactivity can be implemented within modern mapping methods to incorporate additional information and complexity without overloading the user.
Adding too much information to a map may create cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information. 
Complexity and density of visualization methods can overwhelm novice decision makers, however experts are able to use the detail more readily when making-decisions [@cliburn2002design]. 
Supplementary information can be incorporated into online atlases without cluttering the display. Interactive design features found in online cancer maps include tool tip features, drop down menus, data selection, zooming and panning.
Allowing users to explore the map for more information and providing flexibility in the display [@HTLWM]. The use of these supports were found in various online cancer maps identified by @roberts2019communication.
Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction]. Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization). For example, changing the population, age or other demographic variables.
The controls for basic interactive features are often placed outside of the plot space [@TGA], the map image is updated/replaced as the user interacts with the controls. 
Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection interactive capabilities.

Relationships between spatial areas and diseases can be explored with sophistication in nontraditional but still 'cognitively accessible' ways [@TNTEA]. This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution.
The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
The affects of these actions are rapid, incremental, and reversible @DMIV. Where the needs of the audience is changeable and is the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions that allow inspection of the data from many views [@DQBCM]. 
Interactivity and animation provide an alternative to the traditional approach of a single map for a single use, and enable the map to become more exploratory.
Interactivity allows users to drive their own exploration, controlling a display through variable controls, links, tool tips, and data selection tools [@TGA]. Tool tips allow users to obtain additional information regarding specific geographic areas.

Development and adoption of this technology has allowed map users more control of the displayed message. Figure \ref{fig:interacting} presents forms of low level interactions with a user interface, these are the control menus for several online cancer atlases.
The user interface for these online atlases allow combinations of cancer type, gender, and other factors to be shown on the map display and the user to move between and/or compare different views of these variables. 
These controls may also change the geographies shown, such as continents or countries, states or counties. These displays can show familiar shapes for users to orient themselves on the map, then zoom in to explore densely populated communities.
Interactions with online atlases can also change the amount of detail shown.

It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data. @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics. 
Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points. Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
Monmonier implemented linking choropleths with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas. The statistical variation of additional variables can be presented in other appropriate displays.
Low level interaction is more common, this occurs via the user interface or the map.
Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.

@@

@@

Animations can be implemented to allow users varying degrees of control.
Figure \ref{fig:animating} a gives an example of how time can be included through animating displays. This is a basic form of interactivity with animation allowing users to play, pause and rewind films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
Where the message is most important, static or animated graphics allow control over the display and interpretation.
User interaction more allow clicks to progress between stages of more complex animations.
The Australian Cancer Atlas [-@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their own exploration.
